### 2022-04-11 src/text/3-applications-subsections/temporal-analysis.tex
fitting stacked for lc

A common use case in most astrophysical scenarios is to study the temporal variability of the source. The most basic way to do this is to construct a lightcurve, ie, the flux of the source in each given time bin. In gammapy, this is done by using the `LightCurveEstimator` which fits the normalisation of the source in each energy band per observation. For custom time binning, an observation needs to be split into finer ones using the `Observation.select_time` method. Figure \ref{fig:hess_lightcurve_pks} shows the lightcurve of the blazar PKS 2155-304 in different energy bands as observed by the H.E.S.S. telescope during an exceptional flare on the night of July 29 - 30, 2006 \cite{2009A&A...502..749A}. The data is available publicly as a part of the HESS-DL3-DR1 \cite{HESS-DL3-DR1}, and shipped with GAMMAPY_DATA. Each observation is first split into 10 min smaller observations, and spectra extracted for each of these within a 0.11 deg radius around the source. A PowerLawSpectralModel with an index of 3.4 is assumed for all the datasets, and the `LightCurveEstimator` run for two energy bands, 0.5 - 1.5 TeV, and 1.5 - 20 TeV, resp.

Lightcurve in two energy bands (600 GeV - 1.5 TeV, and 1.5 TeV to 20 TeV) for
PKS2155-304 flare seen by H.E.S.S. as in our notebooks.

\begin{figure*}[t]
	\centering
	\includegraphics[width=1.\textwidth]{figures/hess_lightcurve_pks.pdf}
	\caption{10 min binned lightcurve for PKS~2155-304 in two energy bands, (500 GeV - 1.5 TeV, and 1.5 TeV to 20 TeV) as observed by the H.E.S.S. telescopes in 2006.}
	\label{fig:hess_lightcurve_pks} 
\end{figure*}

Instead of extracting a lightcurve, it is also possible to directly fit temporal models to the reduced datasets. By associating an appropriate SkyModel, consisting of both temporal and spectral components, to each dataset, a joint fitting across the datasets will directly return the best fit temporal and spectral parameters.

>>>>

A common use case in most astrophysical scenarios is to study the temporal variability of the source. The most basic way to do this is to construct a lightcurve, ie, the flux of the source in each given time bin. In gammapy, this is done by using the LightCurveEstimator which fits the normalisation of the source in each energy band per observation, keeping other parameters constant. For custom time binning, an observation needs to be split into finer ones using the `Observation.select_time` method. Figure \ref{fig:hess_lightcurve_pks} shows the lightcurve of the blazar PKS 2155$-$304 in different energy bands as observed by the H.E.S.S. telescope during an exceptional flare on the night of July 29 - 30, 2006 \cite{2009A&A...502..749A}. The data is available publicly as a part of the HESS$-$DL3$-$DR1 \cite{HESS-DL3-DR1}, and shipped with GAMMAPY$_$DATA. Each observation is first split into 10 min smaller observations, and spectra extracted for each of these within a 0.11 deg radius around the source. A PowerLawSpectralModel is fit to all the datasets, leading to a reconstructed index of $3.54 \pm 0.02$. With this assumed spectral model the `LightCurveEstimator` run directly for two energy bands, 0.5 - 1.5 TeV, and 1.5 - 20 TeV,  respectively.

\begin{figure*}[t]
	\centering
	\includegraphics[width=1.\textwidth]{figures/hess_lightcurve_pks.pdf}
	\caption{10 min binned lightcurve for PKS~2155$-$304 in two energy bands, (500 GeV - 1.5 TeV, and 1.5 TeV to 20 TeV) as observed by the H.E.S.S. telescopes in 2006.}
	\label{fig:hess_lightcurve_pks} 
\end{figure*}

The obtained flux points can be modelled using any of the available temporal and spectral models, or even using custom temporal models with spectroscopic variability. Alternatively, instead of  extracting a lightcurve, it is also possible to directly fit temporal models to the reduced datasets. By associating an appropriate SkyModel, consisting of both temporal and spectral components, to each dataset, a joint fitting across the datasets will directly return the best fit temporal and spectral parameters.
